---
title: "November 30, 2018 VO Genome Meeting"
author: "Jesse Garcia"
date: "11/26/2018"
output: 
  pdf_document: beamer_presentation
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(warning=FALSE)
library(tidyverse)
library(glue)
library(gghighlight)

theme_set(theme_bw(base_size = 20)) 

theme_set(theme_bw())
theme_update(# axis labels
             axis.title = element_text(size = 30),
             # tick labels
             axis.text = element_text(size = 20),
             # title 
             title = element_text(size = 50))
```



<!-- ## Creating A PSMC Plot with yings 8 resequenced my 22 --> 

```{r cars, include=F,results="hold"}
mu <- 1.01e-8
qlobGen <- 50
qrobGen <- 30


resequenced<-readRDS("../data/SouthernQlobataResequenced_Nov26_2018.rds") %>% mutate(Type="Q. lobata Samples from Nov. 13 I Did")


resequencedMSMC<-"../data/resequenced/"
msmcFiles<-list.files(resequencedMSMC, full.names = T, pattern="final.txt")
id<-str_extract(msmcFiles, pattern="QL.*00F")

msmcFilesDf<-tibble(
  msmcFiles=msmcFiles,
  individuals=id
  
)

msmcFilesDf$files<-msmcFilesDf$msmcFiles %>% map(~read_delim(.x, delim="\t", col_names = T))

msmcFilesDf <-msmcFilesDf %>% unnest(files)



msmcFilesDf<-msmcFilesDf  %>% 
  mutate(Time=left_time_boundary/mu*qlobGen, EffectivePopulationSize= (1/lambda_00)/(2*mu) ) %>% 
  mutate(Type="Q. lobata resequenced genomes Ying Did", species="Q. lobata")


allResequenced<-bind_rows(msmcFilesDf, resequenced)

justQlobata<-allResequenced %>%
  ggplot(aes(x=Time, y=EffectivePopulationSize, colour=individuals, linetype=Type)) +
  geom_step(size=1) +                                                                                                                                                                                  
  theme_bw() +                                                                                                                                              
  scale_x_continuous(trans='log10',labels = scales::comma) +                                                                                                                                   
  scale_y_continuous(trans='log10') + 
  labs(x="Years Ago", y="Effective Population Size")




 #ggsave(filename="../analysis/Qlobata_PSMC_JesseResequence_YingResequenced_Nov28_2018.pdf", plot=justQlobata, height = 10, width=20)



bootstraps<-readRDS("../data/bootstraps200IterationsEmpirical_Nov8.rds")


bootstraps<-bootstraps %>%
  mutate(id=str_extract(files, pattern = "bootstraps_[0-9]+")) %>%
  mutate(species=str_extract(files, pattern="q[lr]ob")) %>%
  mutate(Type=case_when(
    species=="qlob" ~ "Q. lobata reference genome bootstrapped (MaxIterations 100; 100 Bootstraps)",
    species=="qrob" ~ "Q. robur reference genome (MaxIterations 100; 100 Bootstraps)",
    
  )) 

qlobBootstraps<-bootstraps %>% 
  filter(species=="qlob")  %>% 
  mutate(Time=left_time_boundary/mu*qlobGen, EffectivePopulationSize= (1/lambda_00)/(2*mu) ) %>%
  mutate(individuals="Q. lobata reference genome bootstrapped")

qrobBootstraps<-bootstraps %>% 
  filter(species=="qrob")  %>% 
  mutate(Time=left_time_boundary/mu*qrobGen, EffectivePopulationSize= (1/lambda_00)/(2*mu) ) %>%
  mutate(individuals="Q. robur reference genome bootstrapped")

allPSMC<-allResequenced %>%
  ggplot(aes(x=Time, y=EffectivePopulationSize,colour=Type, group=interaction(individuals, Type))) +
  geom_step() +
  geom_step(data=qlobBootstraps, aes(x=Time, y=EffectivePopulationSize, group=id)) +
  geom_step(data=qrobBootstraps, aes(x=Time, y=EffectivePopulationSize, group=id)) +
  theme(text = element_text(size=16)) +
  theme_bw() +                                                                                                                                              
  scale_x_continuous(trans='log10',labels = scales::comma) +                                                                                                                                   
  scale_y_continuous(trans='log10') + 
  labs(x="Years Ago", y="Effective Population Size")
 
 
 
#ggsave(filename="../analysis/QlobataAndQRobur_PSMC_JesseResequence_YingResequenced_Nov28_2018.pdf", plot=allPSMC, height = 10, width=20)
```



## MSMC Plot with Ying's 8 resequenced oak genomes and 22 oak genomes I processed
```{r}
 allPSMC +
    theme(text = element_text(size=18), legend.position="bottom")+ guides(colour = guide_legend(ncol = 1))
```



## Visually it looks like there are 2 outliers among the genomes I processed
```{r}
justQlobata<-allResequenced %>% 
  filter(!is.na(files)) %>%
  ggplot(aes(x=Time, y=EffectivePopulationSize, colour=individuals)) +
  geom_step(size=1) +                                                                                                                                                                                  
  theme_bw() +                                                                                                                                              
  scale_x_continuous(trans='log10',labels = scales::comma) +                                                                                                                                   
  scale_y_continuous(trans='log10') + 
  labs(x="Years Ago", y="Effective Population Size") + 
  gghighlight(individuals %in% c("QL.BER.1", "QL.HV.1"))
  

justQlobata  +
    theme(text = element_text(size=18))
```


##
\begin{center}
How might coverage affect demographic inference?
\end{center}




## No strong correlation between population scaled mutation rate (~ Heterozygosity) and mean coverage

```{r pressure, echo=FALSE}
files<-"../data/MSMCMutationRate_Coverage_Df_November28.rds"
mutationRateMSMC<-readRDS(files)

mutationRateMSMC %>% 
  ggplot(aes(x=mean, y=mutationRate, colour=individuals)) +
  geom_point() + 
  geom_smooth() +
  labs(x="Mean Coverage (MQ > 20)", y="Heterozygosity (Segregating Sites/Callable Sites/Waterson/2)") +
  theme_bw()  +
    theme(text = element_text(size=14))




```

## The two visual outliers in Heterozygosity are also the outliers at Time 0

```{r}
mutationRateMSMC %>% 
  ggplot(aes(x=mean, y=mutationRate, colour=individuals)) +
  geom_point() + 
  geom_smooth() +
  labs(x="Mean Coverage (MQ > 20)", y="Heterozygosity (Segregating Sites/Callable Sites/Waterson/2)") +
  theme_bw() + 
  gghighlight(individuals %in% c("/QL.BER.1.00F", "/QL.HV.1.00F")) +
    theme(text = element_text(size=18))

```




## No correlation of effective population size at Time 0 and Mean Coverage

```{r}
#allResequenced %>% filter(Time==0) %>% arrange(desc(EffectivePopulationSize))

test=allResequenced %>% dplyr::mutate(msmcFiles=str_replace(msmcFiles, pattern="../data/resequenced/", replacement = "")) %>%
  mutate(individuals=str_replace(msmcFiles, pattern=".final.txt", "")) %>% filter(Time==0)

test=test %>%
  mutate(files=str_replace(files, "/u/flashscratch/j/jessegar/FilteringResequenceHighP", replacement = "")) %>%
  mutate(files=str_replace(files, pattern = ".final.txt", replacement=""))

test<-test %>% mutate(individuals=coalesce(individuals, files))

test<-left_join(test,mutationRateMSMC)



test %>% 
  ggplot(aes(x=mean,y=EffectivePopulationSize, colour=individuals)) +
  geom_point() + 
  theme_bw() + 
  labs(x="Mean Coverage (MQ > 20)", y="Effective Population Size At Time 0") +
  theme_bw() + 
  gghighlight(individuals %in% c("/QL.BER.1.00F", "/QL.HV.1.00F")) +
    theme(text = element_text(size=18))
```


##  Heterozygosity and Population Size at Time 0 does appear to be moderately correlated. Ying's samples aren't on this plot but I think they would strengthen this correlation. 
```{r}
test=allResequenced %>% dplyr::mutate(msmcFiles=str_replace(msmcFiles, pattern="../data/resequenced/", replacement = "")) %>%
  mutate(individuals=str_replace(msmcFiles, pattern=".final.txt", "")) %>% filter(Time==0)

test=test %>%
  mutate(files=str_replace(files, "/u/flashscratch/j/jessegar/FilteringResequenceHighP", replacement = "")) %>%
  mutate(files=str_replace(files, pattern = ".final.txt", replacement=""))

test<-test %>% mutate(individuals=coalesce(individuals, files))

test<-left_join(test,mutationRateMSMC) %>% filter(Time==0)



test %>% 
  ggplot(aes(x=mutationRate,y=EffectivePopulationSize, colour=individuals)) +
  geom_point() + 
  theme_bw() + 
  labs(x="Heterozygosity (Segregating Sites/Callable Sites/Waterson/2)", y="Effective Population Size At Time 0") +
  theme_bw() + 
  gghighlight(individuals %in% c("/QL.BER.1.00F", "/QL.HV.1.00F")) +
    theme(text = element_text(size=18))
```







## Mean effective population size across all of Time and Mean Coverage doesn't appear correlated
```{r}

test=allResequenced %>% dplyr::mutate(msmcFiles=str_replace(msmcFiles, pattern="../data/resequenced/", replacement = "")) %>%
  mutate(individuals=str_replace(msmcFiles, pattern=".final.txt", ""))


test=test %>%
  mutate(files=str_replace(files, "/u/flashscratch/j/jessegar/FilteringResequenceHighP", replacement = "")) %>%
  mutate(files=str_replace(files, pattern = ".final.txt", replacement=""))

test<-test %>% mutate(individuals=coalesce(individuals, files))

test<-left_join(test,mutationRateMSMC)

test %>% 
  group_by(individuals, mutationRate, mean) %>% 
  summarise(meanNe=mean(EffectivePopulationSize)) %>%
  ggplot(aes(x=mean,y=meanNe, colour=individuals)) +
  geom_point() + 
  theme_bw() + 
  labs(x="Mean Coverage (MQ > 20)", y="Mean Effective Population Size Across Time") +
  theme_bw() + 
  gghighlight(individuals %in% c("/QL.BER.1.00F", "/QL.HV.1.00F")) +
    theme(text = element_text(size=18))



```



## Pi also suggests that these two samples have higher levels of heterozygosity on average than other samples
```{r}
piPerSite<-readRDS("../data/PiPerSiteResequencedGenomes_November292018.rds")


piPerSite %>% ungroup() %>%
  rename(start=X1, end=X2, totalCallable=X3, pi=X4) %>% 
  mutate(piPerSite=pi/totalCallable) %>% 
  ggplot(aes(x=individuals, y=piPerSite, colour=individuals)) + 
  geom_boxplot() + 
  theme_bw() + 
  gghighlight(individuals %in% c("QL.BER.1.00F.2018", "QL.HV.1.00F.2018")) + 
  labs(x="Individuals", y="Pi per callable site in 500kb Windows", colour="Individuals") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    theme(text = element_text(size=18))

```


## Questions

* Are the samples I recently recieved from Sorel more closely related to the reference genome than the one's Ying processed? 